### 2012-02-24 papers/eccv2012/main.tex
sections 2 and 3 written

\begin{abstract}
  We consdier the problem of learning to reconstruct polygonal models
  from single and multiple views.

  In this difficult domain we start small and focus on indoor
  Manhattan models.

  We learn lots about practical structured prediction in this unique
  domain

  We kick the state-of-the-art's butt on everything :)
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Reconstruction long considered a central problem in computer vision

Multiple view reconstruction long history, mostly uses probabilistic
models with few free parameters

Single view reconstruction a recent but now well-established
trend. Learning obviously a core part of this fundamentally
under-determined problem. But not much care on loss functions,
optimality guarantees, or statistical consistency.
-> Exceptions: Hedau et al (but very simple models)

Structured prediction used for increasingly complex models. Its use in
computer vision part of a long trend towards statistically rigorous
and well-understood convex optimization techniques for solving
empirical risk minimization problems in computer vision.

Has been used much before for reconstruction. Stereo for example, and
Hedau. Here we go all the way to an expressive high-level geometric
model class.

Attraction of doing this is many fold:
Can minimize model-level loss rather than pixel-level loss
-> give a nice demonstration of this
Can minimize realistic loss functions in a rigorous way
Can learn in both single and multiple view setting using the *same*
model
Bayesian inference algorithm

We learn much about the practice of structured prediction of value to
the community and focus an entire section of conveying these lessons.

Out approach: describe the hypothesis class, describe the model

The remainder of this paper is...

\section{Background}
\label{sec:background}

We're drawing together multiple view reconstruction, single view
reconstruction, and structured prediction -> cannot give a reasonable
survey, sorry!!

Reconstruction papers

Multi view reconstruction, stereo -> not much learning

Single view -> learning but mostly at pixel level. Some exceptions

Learning papers

Structured prediction, reconstruction applications

\section{Model}
\label{sec:model}

We now describe the indoor Manhattan model that we are attempting to
learn. As with any learning problem, there are three ingredients that
we need to specify: a hypothesis class, an observation model for
hypotheses, and a loss function. In our setup these are, respectively,
the class of indoor Manhattan models, a log--linear Bayesian
likelihood, and either the relative depth error or a labelling error
(we describe both).

\subsection{Hypothesis Class}

This paper is concerned with the hypothesis class consisting of indoor
Manhattan reconstructions, which are 3D polygonal models characterized
by infinite floor and ceiling planes with vertical walls extending
between them \cite{Lee09,Flint10cvpr}. Indoor Manhattan environments
are a sub--class of general Manhattan environments. Examples are shown
in figure XXX.

This is an attractive hypothesis class because
\begin{enumerate}
\item{it captures many \textit{regularities} within man--made
  environments;}
\item{the geometric primitives (floor/wall/ceiling) are immediately
  useful for semantic--level scene understanding;}
\item{it is expressive enough to represent approximately or exactly
  a surprisingly wide variety of environments;}
\item{there is a simple and convenient parametrisation;}
\item{an efficient inference algorithm exists \cite{Flint10eccv}.}
\end{enumerate}

An indoor Manhattan model is much like an archicetural floor--plan,
and can be specificied as a set of 2D line segments representing walls
together with the position of the floor and ceiling plane (figure
XXX). In this paper we adopt an image--domain parametrisation due to
its convenience for inference and learning. Following Flint \etal
\cite{Flint11} we represent hypotheses as a path running from the left
edge to the right edge of the image, which we will refer to as a
\textit{seam}. A seam $\Seam$ consists of a sequence of pairs of
scalars,
\begin{equation}
  \Seam = \{\seam_i,\Label_i\}_{i=1}^{\Width}
\end{equation}
where $\seam_i$ is the $y$--cordinate at which the path intersects
image column $i$, $\Orient_i$ is the orientation of the wall in that
column, and $\Width$ is the image width. Remarkably, this simple
parametrisation specifies a unique metric 3D model up to scale
\cite{Flint11}.

While we do not here have space for a full discussion of the geometry
of indoor Manhattan models, there are two properties of the seam
representation that will be of relevance to the following
sections. Firstly, images of indoor Manhattan environments can be
rectified so that vertical lines in the world project to vertical
lines in the image \cite{Flint10eccv}. Secondly, indoor Manhattan
models decompose in the following manner. We said earlier that a seam
$\Seam$ specifies a unique 3D reconstruction. Let
$\Orient(x,y;\Seam)\in\TheOrientations$ be the orientation of the 3D
surface projecting to pixel $(x,y)$. Then Flint \etal
\cite{Flint10eccv} showed that $\Orient$ is functionally dependant
\textit{only} on the labels for column $x$. That is, we may write
\begin{equation}
  \label{eq:orient-indep}
  \Orient(x,y;\Seam) = \tilde{\Orient}(x,y;\seam_x,\Label_x) ~.
\end{equation}
Similarly, letting $\Depth$ be the depth of the reconstruction at
$(x,y)$ we may write $\Depth$ in terms of a single label pair,
\cite{Flint10iccv}
\begin{equation}
  \label{eq:depth-indep}
  \Depth(x,y;\Seam) = \tilde{\Depth}(x,y;\seam_x,\Label_x) ~.
\end{equation}

\subsection{Observation Model}

We adopt the probabilistic model formulated by Flint \etal
\cite{Flint11}, which relates indoor Manhattan hypotheses to
single--view image features, multiple--view photoconsistency terms,
and a reconstructed point cloud. From our perspective these are simply
features and any subset may be omitted if in appropriate for a given
application.

TODO: walk through the graphical model?

Under this model the prior on environments is a geometric distribution
in the number of distinct wall segments, which means, roughly, that we
consider reconstructions with more walls apriori less likely. The
benefit of this prior is that its logarithm is linear in the number of
wall segments,
\begin{equation}
  \log P(\Seam) = - \vect{n} \cdot \Penalties
\end{equation}
where $\vect{n}$ is a vector containing the number of walls in $\Seam$
of various categories (concave, convex, occluding) and $\Penalties$ is
a vector of hyper--parameters.

Flint \etal showed that for a variety of single--view and
multiple--view features $\Features$ there are reasonable choices of
likelihood that can be written in the form
\begin{equation}
  \log P(\Features ~|~ \Seam) = \sum_{x=1}^\Width 
  \PixelPayoff_{\Features}(x,\seam_x)
\end{equation}
where $\PixelPayoff_{\Features}$ is a 2--dimensional matrix computed
deterministically from $\Features$ \cite{Flint11}. It is easy to show
that for each likelihood in \cite{Flint11}, the log--likelihoods
is linear in the hyper--parameters $\LikelihoodHypers$,
\begin{equation}
  \log P(\Features ~|~ \Seam) = 
  \sum_{x=1}^\Width
    \LikelihoodHypers_{\Features} 
    \cdot
    \PixelFtr_{\Features}(x,\seam_x)
\end{equation}

Assuming conditional independence between the feature types, we have
for the posterior,
\begin{eqnarray}
  P(\Seam ~|~ \Features_1 \ldots \Features_n)
  &\propto&
  P(\Seam) P(\Feature_1 ~|~ \Seam) \ldots P(\Feature_n ~|~ \Seam)\\
  \log P(\Seam ~|~ \Features_1 \ldots \Features_n)
  &=&
  - \vect{n} \cdot \Penalties
  + \sum_{i=1}^n \sum_{x=1}^\Width 
      \LikelihoodHypers_i
      \cdot
      \PixelFtr_i(x,\seam_x)
  + O(1) ~.
  \label{eq:log-posterior}
\end{eqnarray}

In order to place our model within a structured prediction framework
we need to re--write the above in terms of a joint feature space
$\JointFtr(\Seam,\Features)$. The decomposability of indoor Manhattan
mdoels into payoff matrices permits precisely such a
formulation. Defining
\begin{align}
  \JointFtr(\Seam,\Features) =
  \begin{bmatrix}
    -\vect{n} \\
    \sum_{x=1}^\Width \PixelFtr_1(x,\seam_x) \\
    \vdots \\
    \sum_{x=1}^\Width \PixelFtr_n(x,\seam_x)
  \end{bmatrix}
  & \quad &
  \Model =
  \begin{bmatrix}
    \Penalties \\
    \LikelihoodHypers_1 \\
    \vdots \\
    \LikelihoodHypers_n
  \end{bmatrix}
  \label{eq:joint-ftr-def}
\end{align}
we see that \eqnref{log-posterior} can be written
\begin{equation}
  \log P(\Seam ~|~ \Features_1 \ldots \Features_n)
  =
  \bigl\langle \Model ~,~ \JointFtr(\Seam,\Features) \bigr\rangle + O(1)
\end{equation}
where we have adopted Hilbert space notation with
$\langle\cdot,\cdot\rangle$ denoting an inner product. Since $\Model$
contains all the free parameters in the model, the goal learning will
be to optimise $\Model$ with respect to a set of training features
$\{\Features_i\}$ together with ground truth labellings $\{\Seam_i\}$.

\subsubsection{Features}

The precise make--up of the feature space depends on the particular
forms for the likelihoods and which sensor modalities are
available. We define separate feature spaces for the single-- and
multiple--view contexts; both are summarized in \figref{featurespace}.

\begin{figure}[tb]
  \centering
  \begin{tabular}{@{}llll@{}}
    \toprule
    Feature & Dimensionality & Single--view? & Reference \\
    \midrule
    Line sweeps & 1 & yes & Lee \etal \cite{Lee09}\\
    Stereo photo--consistency & 4 & no & Flint \etal \cite{Flint11}\\
    Point cloud & 2 & no & Flint \etal \cite{Flint11}\\
    \bottomrule
  \end{tabular}
  \caption{The composition of our feature space. Each row corresponds
    to one or more of the $\PixelFtr_i$ in \eqnref{joint-ftr-def}.}
  \label{fig:featurespace}
\end{figure}

\subsection{Loss Functions}

A loss function $\Delta(\Seam,\EstSeam)$ is a measure of the
deviation of a predicted label $\EstSeam$ from the ground truth
$\Seam$. In the context of learning one often faces a trade--off
between choosing a loss that leads to tractible optimization, and
choosing a loss that measures what one ``really'' cares about. For
example, Hoiem \etal \cite{Hoiem05} implicitly use a per--pixel loss,
since this decomposes optimization pixel--wise. However, it would be
hard to argue that pixel--wise accuracy is an honest measure of what
the authors care about: some pixel--level errors are insignificant to
the overall quality of a reconstruction, while others are
catastrophic. Whether the authors care about the accuracy of their
reconstruction in 3D, its utility for contextual object detection, or
even the aesthetic appeal of reconstructions, a per--pixel loss is
almost certainly a weak measure of success.

A more reasonable choice would be the relative depth error, which has
been the gold standard within the reconstruction community for decades
\cite{Hartley04}, and measures the average deviation between
reconstructed and ground truth depths:
\begin{equation}
  \DepthLoss(\Seam,\EstSeam)
  =
  \frac{1}{N}
  \sum_{\Pixel}
    \frac
        {\Depth(\Pixel;\Seam) - \Depth(\Pixel;\EstSeam)}
        {\Depth(\Pixel;\Seam)} ~,
\end{equation}
where $N$ is the number of pixels. Another reasonable choice is the
labelling error, used widely within the semantic segmentation
literature,
\begin{equation}
  \LblLoss(\Seam,\EstSeam)
  =
  \frac{1}{N}
  \sum_{\Pixel}
    \bigl[~\Orient(\Pixel;\Seam) \neq \Orient(\Pixel;\EstSeam)~\bigr] ~,
\end{equation}
where $[p]$ is 1 if $p$ is true and 0 otherwise. An attractive
characteristic of the indoor Manhattan class is that both of these
loss functions can be optimized exactly. The algorithmic details for
this are left to \sectref{learning}; the key result that we need to
attend to here is that $\DepthLoss$ and $\LblLoss$ can be written in a
form similar to the payoff formulation for the feature
likelihoods. Recall the functional independence established in
\eqnref{func-indep}.

In the
case of reconstruction, the gold standard is the relative depth error
\cite{Hartley04}

Should we show how to write these as payoff matrices here, or leave
this to the learning section?

L1, L2 error: easy representation, not meaningful, give examples

Labelling error
-> showing these will require some of the ICCV framework to have been presented
-> give nice intuition, show scatter plots, examples

Relative depth error
-> showing these will require some of the ICCV framework to have been presented
-> give nice intuition, show scatter plots, examples

None of these are pixel classifiers!

Writing loss functions in the seam representation -> maybe just
mention that this will be useful later

\section{Inference}
\label{sec:inference}

Describe the inference algorithm in terms of payoff/penalty
optimization, cite ECCV -> this section must be short!

\section{Learning}
\label{sec:learning}

Risk minimization formulation
%  -> min_f E[ Loss(f(x),y*) ]
Empirical risk problem
%  -> min_f sum_{xi} Loss(f(x),y*)
Write f as a linear argmax
%  -> f(x) = argmax_y <w,psi(x,y)>
Reduced to quadratic optimization (cite SVMstruct papers)
%  -> min_f ...
-> subject to ....
Minimize using algorithm in ...
This requires a separation oracle, which is the topic of the
following subsection

Loss-augmented inference
-> the problem
-> writing loss the functions as payoff matrices
-> writing a payoff matrix for loss-augmented inference
-> write a proof that ``the optimal model payoff matrix
B=... solves the loss-augmented inference problem <w,psi> + L

\section{Practical Considerations}
\label{sec:practical-considerations}

- negative slacks are more harmful than one would expect

- inconsistent inference is very harmful

- inference when the ground truth not in the hypothesis class

- conditioning in a joint feature space

\section{Multiple View Results}
\label{sec:mv-results}

\section{Single View Results}
\label{sec:sv-results}

\section{Conclusion}
\label{sec:conclusion}

Wow, we really did rock didn't we?

Hypothesis class, why it's interesting

Statistical rigour is guuuud

These fields are ready for this kind of unification

Rigorous joint learning for contextual object detection: a challenge

\bibliographystyle{splncs}
\bibliography{AVLstrings,VisionRefs}