### 2014-07-24 iccv2011/rebuttal.txt
move papers/* to top level

We thank all reviewers for encouraging and constructive reviews. We are pleased that all reviewers offered broad encouragement of our approach and noted general satisfaction with our results. We are confident that we can address all suggestions and clarification requests in the camera-ready manuscript.

R1 requested several clarifications, each of which we have addressed in the final paper; we summarise the answers below:
1) Regarding occlusions- we assume locally convex environments (lines 408-417); this is violated surprisingly rarely since we are always interested in the boundaries of the environment and never the objects or clutter within. This means the DP optimization simply ignores occlusions. When we do encounter non-convex environments we find that global optimization together with a prior on model complexity (eq. 4) produces pleasing results regardless.
2) Features- we perform intensity and contrast normalization but we do not handle specularities or other non-linear lighting phenomena. Our intention is to show that models from stereo vision fit naturally into our framework; fully leveraging the best modern stereo algorithms remains future work.
3) Training- We trained and evaluated on separate sets of sequences. No place or room observed in the training set was also present in the test set.
4) Inference- All our experiments were performed with 2 auxiliary frames, which were sampled 1 second before and after each base frame. In the video sequence we selected base frames at 8-second intervals, and at each step we discarded the previous model and re-estimated. Each video frame shows the most recent model re-projected into the current frame.
5) Timing- We include further timing results in the final paper in line with the request by R2. Training on 150 frames takes roughly 15 minutes.

We thank R2 for two suggested citations. The work in (Lee et al, 2010) is relevant to ours primarily via earlier work (Lee et al, CVPR 2009), which our paper cited. We have added a citation for (Hedau et al, 2009), though our paper overlaps only marginally with this monocular approach. We regret that R2 found Section 3 unclear and we will revise extensively for the final submission. We have added timing results to the final paper as requested; we regret that this was omitted from the review manuscript.

We agree with R3 regarding overlap with [7]. We intended this paper to be understandable on its own since [7] was published less than a year ago and is not widely known, so we compromised by repeating some of their results during lines 530-553. We were aware of Erick Delage's early work in this area and regret omitting that citation; we have rectified this in our final submission. Regarding comparison with [7], the "Monocular only" line in Figure 8 is identical to the system of [7]; our final paper states this explicitly. (Compared to [7], the addition of stereo and 3D features reduces mean depth error by 10%.)